Integrin β4 promotes DNA damage drug resistance in triple-negative breast cancer via TNFAIP2/IQGAP1/Rac1

Version 1

Abstract

Summary

Anti-tumor drug resistance is a challenge for triple-negative breast cancer treatment. Our previous work demonstrated that TNFAIP2 activates Rac1 to promote triple-negative breast cancer cell proliferation and migration. However, the mechanism by which TNFAIP2 activates Rac1 is unknown. In this study, we found that TNFAIP2 interacts with IQGAP1 and Integrin β4. Integrin β4 activates Rac1 through TNFAIP2 and IQGAP1 and confers DNA damage drug resistance in triple-negative breast cancer. These results indicate that the Integrin β4/TNFAIP2/IQGAP1/Rac1 axis provides potential therapeutic targets to overcome DNA damage drug resistance in triple-negative breast cancer.

Introduction

Breast cancer is the most common malignant tumor in women, and the morbidity and mortality rates of female malignant tumors are ranked first in the world1. Although the diagnosis and treatment of breast cancer has entered the era of molecular typing,35% of breast cancers still experience recurrence, metastasis and treatment failure2. According to the expression of estrogen receptor (ERα), progesterone receptor (PR) and human epidermal growth factor receptor (HER2), breast cancer is divided into ER/PR-positive, HER2-positive and triple-negative breast cancer (TNBC)3.For ER/PR-positive and HER2-positive breast cancer, endocrine therapies such as tamoxifen and anti-HER2 targeted therapy such as trastuzumab have achieved good efficacy. Targeted drugs for TNBC patients with BRCA1/2 mutations include two PARP inhibitors, olaparib and talazoparib. These targeted drugs cannot fully meet the clinical needs of patients with various TNBC subtypes4. Currently, DNA damage chemotherapy drugs, including epirubicin and cisplatin, are widely used for TNBC treatment.

TNBC often recurs and metastasizes due to the development of chemoresistance, although it is initially responsive to chemotherapeutic drugs5. Chemoresistance severely impacts the clinical outcomes of patients. Tumor cells become resistant to chemotherapeutic agents through several mechanisms, such as improving DNA damage repair, changing the intracellular accumulation of drugs, or increasing anti-apoptotic mechanisms6. Therefore, characterization of the underlying molecular mechanisms by which resistance occurs will provide opportunities to develop precise therapies to enhance the efficacy of standard chemotherapy regimens 7, 8.

TNFAIP2 is abnormally highly expressed in a variety of tumor cells, including TNBC9, nasopharyngeal carcinoma 10, malignant glioma11, uroepithelial carcinoma12and esophageal squamous cell carcinoma13, and is associated with poor prognosis. Our previous work 9,14showed that TNFAIP2, as a KLF5 downstream target protein, can interact with Rac115, a member of the Rho small GTP enzyme family, and activate Rac1 to alter the cytoskeleton, thereby inducing filopodia and lamellipodia formation and promoting the adhesion, migration and invasion of TNBC cells. After activation, Rac1 can activate AKT, PAKs, NADPH oxidase and other related signaling pathways to promote cell survival, proliferation, adhesion, migration and invasion16. Activation of Rac1 can reduce the therapeutic response to trastuzumab in breast cancer and increase the resistance of TNBC cells to paclitaxel17, but the specific mechanism of action is not completely clear.

Rac1 has been shown to play an important role in DNA damage repair. Activated Rac1 can promote the phosphorylation of the DNA damage response-related molecules ATM/ATR, CHK1/2 and H2AX by activating the activity of protein kinases such as ERK1/2, JNK and p38 18, 19, thus improving the DNA damage repair ability and inhibiting tumor cell apoptosis20–22. Rac1 also promotes aldolase release and activation by changing the cytoskeleton and activates the ERK pathway to increase thepentose phosphate pathway to promote nucleic acid synthesis, providing more raw materials for DNA damage repair23, 24. At the same time, the interaction of Rac1 and PI3K promoted AKT phosphorylation and glucose uptake25, 26. Therefore, Rac1 is well established to promote the chemoresistance of breast cancer by promoting DNA damage repair.

Integrin β4 (ITGB4) is a major component of hemidesmosomes and a receptor molecule of laminin. Studies have shown that laminin-5 interacts with ITGB4 to activate Rac1 activity and promote cell migration27 and polarization28 by altering the cytoskeleton. Since ITGB4-positive cancer stem cell (CSC)-enriched mesenchymal cells were found to reside in an intermediate epithelial/mesenchymal phenotypic state, ITGB4 can be used to enable stratification of mesenchymal-like TNBC cells29. In addition, the expression of ITGB4 on ALDHhigh breast cancer and head and neck cancer cells was significantly greater than that on ALDHlow cells, proving the effects that ITGB4 targets on both bulk and CSC populations30. Furthermore, ITGB4-overexpressing TNBC cells provided cancer-associated fibroblasts (CAFs) with ITGB4 proteins via exosomes, and ITGB4-overexpressing CAF-conditioned medium promoted the proliferation, epithelial-to-mesenchymal transition, and invasion of breast cancer cells31. ITGB4 also promotes breast cancer cell resistance to tamoxifen-induced apoptosis by activating the PI3K/AKT signaling pathway and promotes breast cancer cell resistance to anoikis by activating Rac132. However, how ITGB4 activates Rac1 is not completely clear.

Rac1 activity is regulated by guanylate exchange factors (GEFs), GTPase activation proteins (GAPs), and guanine separation inhibitors (GDIs)33. GAPs typically provide the necessary catalytic groups for GTP hydrolysis, but not all GAPs function as hydrolases. IQGAP1 lacks an arginine in the GTPase binding domain and cannot exert the hydrolysis effect of GAPs34. IQGAP1 can increase the activity of Rac1 and CDC4235, 36.

In this study, we demonstrated that TNFAIP2 interacts with IQGAP1 and ITGB4. ITGB4 promotes TNBC drug resistance via the TNFAIP2/IQGAP1/Rac1 axis by promoting DNA damage repair. Our results suggest that ITGB4 and TNFAIP2 might serve as promising therapeutic targets for TNBC.

Results

TNFAIP2 promotes TNBC DNA damage drug resistance

To explore the functional significance of TNFAIP2 in TNBC drug resistance, we constructed stable TNFAIP2 overexpression and TNFAIP2 knockdown HCC1806 and HCC1937 cells. As shown in Figure 1A-E, overexpression of TNFAIP2 significantly increased cell viability when treated with EPI and BMN. Additionally, knockdown of TNFAIP2 significantly decreased cell viability when treated with EPI and BMN (Fig. 1F-J). We then examined the effects of TNFAIP2 on DNA damage repair and found that TNFAIP2 promotes DNA damage repair in response to EPI and BMN. TNFAIP2 overexpression decreased the protein expression levels of γH2AX, a marker of DNA damage, and cleaved-PARP, a marker of apoptosis (Fig. 1K). Additionally, knockdown of TNFAIP2 significantly increased γH2AX and cleaved-PARP protein expression levels in response to EPI and BMN in both cell lines (Fig. 1L).

TNFAIP2 promotes TNBC DNA damage drug resistance

(A-E) Stable TNFAIP2 overexpression in HCC1806 and HCC1937 cells significantly increased cell viability in the presence of EPI (0-1.6 μM) or BMN (0-40 μM) treatment for 48h, as measured by the SRB assay. TNFAIP2 protein expression was detected by WB.

(F-J) Stable TNFAIP2 knockdown in HCC1806 and HCC1937 cells significantly decreased cell viability in the presence of EPI (0-1.6 μM) or BMN (0-40 μM) treatment for 48h, as measured by the SRB assay. TNFAIP2 protein expression was detected by WB.

(K) TNFAIP2 promoted DNA damage repair in the presence of EPI and BMN. HCC1806 and HCC1937 cells stably overexpressing TNFAIP2 were treated with 400 or 800 nM EPI for 48h and 10 μM BMN for 24h, respectively. TNFAIP2, γH2AX, and PARP protein expression was detected by WB.

(L) TNFAIP2 knockdown increased DNA damage in the presence of EPI and BMN. Stable TNFAIP2 knockdown cells were treated with 400 or 800 nM EPI for 24 or 48h and 2.5 μM BMN for 24h. TNFAIP2, γH2AX, and PARP protein expression was detected by WB.

The function of TNFAIP2 was further validated by using two other DNA damage drugs, DDP and AZD (Fig. S1A-L). These results suggested that TNFAIP2 enhances TNBC cell drug resistance by promoting DNA damage repair.

TNFAIP2 confers TNBC drug resistance in vivo

To test whether TNFAIP2 also decreases the sensitivity of TNBC cells to EPI and BMN in vivo, we performed animal experiments in nude mice. HCC1806 cells with stable TNFAIP2 knockdown were orthotopically inoculated into the fat pad of 7-week-old female mice (n=8 or 12/group). When the tumor mass reached approximately 50 mm3, each group was divided into two subgroups to receive either EPI (2.5 mg/kg, twice a week) or vehicle control for 23 days and either BMN (1 mg/kg, twice a week) or vehicle control for 29 days. We observed that depletion of TNFAIP2 suppressed breast cancer cell growth in vivo. This is consistent with our previous report9. More importantly, TNFAIP2 depletion further decreased tumor volume when mice were treated with EPI and BMN (Fig. 2A-F). Meanwhile, BMN treatment had no effect on the body weight of mice (Fig. S2F). Consistently, EPI and DDP generated similar results but decreased mouse body weight due to their high toxicity (Fig.S2D-E). These results suggest that inhibition of TNFAIP2 expression can overcome HCC1806 breast cancer cell drug resistance in animals.

TNFAIP2 confers TNBC drug resistance in vivo

(A-F) TNFAIP2 knockdown increased the sensitivity of HCC1806 breast cancer cells to EPI and BMN in vivo. HCC1806 cells with stable TNFAIP2 knockdown were transplanted into the fat pad of 7-week-old female nude mice. When the average tumor size reached approximately 50 mm3 after inoculation, mice in each group were randomly divided into two subgroups (n=4/group) to receive EPI (2.5 mg/kg), BMN (1 mg/kg) or vehicle control for 23 or 29 days (A-B). Tumor size was measured twice a week (C-D), and tumor masses were collected and weighed at the end of the experiments (E-F). *: P<0.05, **: P<0.01, t-test.

TNFAIP2 promotes TNBC drug resistance and DNA damage repair via Rac1

Since chemotherapeutic agents and PRAP inhibitors induce DNA damage directly or indirectly, DNA damage repair ability profoundly affects the sensitivity of cancer cells to these therapies37, 38. Since TNFAIP2 can activate Rac1, a well-known drug resistance protein, we investigated whether TNFAIP2 induces chemotherapeutic resistance through Rac1. We found that Rac1 knockdown rescued the effects of TNFAIP2 overexpression-induced drug resistance to EPI and BMN in HCC1806 and HCC1937 cells (Fig. 3A-F). We also found that γH2AX and cleaved-PARP protein levels were up-regulated again in Rac1 knockdown and TNFAIP2-overexpressing HCC1806 and HCC1937 cells in response to EPI and BMN (Fig. 3G-J). We obtained similar results by using DDP and AZD treatment (Fig. 3SA-J). Collectively, these results suggest that TNFAIP2 promotes DNA damage repair and drug resistance via Rac1.

TNFAIP2 promotes TNBC drug resistance and DNA damage repair via Rac1

(A-F) Rac1 knockdown abolished TNFAIP2-induced TNBC resistance to EPI and BMN. HCC1806 (A-C) and HCC1937 (D-F) cells with stable TNFAIP2 overexpression were transfected with Rac1 or control siRNA, followed by treatment with EPI (0-1600 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Protein expression levels were analyzed by WB.

(G-J) Rac1 depletion abolished TNFAIP2-induced DNA damage decrease in response to EPI and BMN. HCC1806 (G-H) and HCC1937 (I-J) cells with stable TNFAIP2 overexpression were transfected with Rac1 or control siRNA, followed by treatment with EPI (400 or 800 nM) and BMN (10 μM) for 24h, respectively. Protein expression levels were analyzed by WB.

IQGAP1 mediates Rac1 activation by TNFAIP2 and promotes TNBC drug resistance

To characterize the mechanism by which TNFAIP2 activates Rac1, we performed an IP-MS experiment. We found that TNFAIP2 interacts with IQGAP1 (Fig. 4A). To validate whether TNFAIP2 interacts with IQGAP1, we constructed HCC1806 cells with stable Flag-TNFAIP2 overexpression and collected Flag-tagged TNFAIP2 cell lysates for immunoprecipitation assays using Flag-M2 beads (Fig. S4A). We performed immunoprecipitation using ananti-IQGAP1 antibody and found that endogenous IQGAP1 protein interacted with endogenous TNFAIP2 protein inHCC1806 cells (Fig. 4B). Next, we mapped the regions of TNFAIP2 and IQGAP1 proteins responsible for the interaction by generating a series of Flag-TNFAIP2 deletion mutants and transfected them into HEK293T cells together with full-length IQGAP1. Then, we performed immunoprecipitation assays using Flag-M2 beads (Figure S4B), and the N-terminus (1-79 aa) of the TNFAIP2 protein interacted with IQGAP1. To explore the function of IQGAP1 in TNBC drug resistance, we knocked down IQGAP1 in HCC1806 and HCC1937 cells. As shown in Figure 4C-G, knockdown of IQGAP1 significantly decreased cell viability in the presence of EPI and BMN in both cell lines. We also examined the effects of IQGAP1 on DNA damage repair and found that IQGAP1 promotes DNA damage repair. IQGAP1 knockdown increased γH2AX and cleaved-PARP protein expression levels when HCC1806 and HCC1937 cells were treated with EPI and BMN (Fig. 4H). Next, we found that IQGAP1 knockdown abrogated the effects of TNFAIP2 overexpression on resistance to EPI and BMN (Fig. 4I-K, Fig.S4C-E). We also found that γH2AX and cleaved-PARP protein levels were up-regulated in IQGAP1 knockdown and TNFAIP2-overexpressing HCC1806 and HCC1937 cells (Fig. 4L). In addition, we found that the TNFAIP2 overexpression-induced increase in Rac1 activity was abolished by IQGAP1 knockdown (Fig. 4M).

IQGAP1 mediates Rac1 activation by TNFAIP2 and promotes TNBC drug resistance

(A) The IP-MS result of TNFAIP2 in HCC1806 cells.

(B) Endogenous TNFAIP2 interacts with IQGAP1 in HCC1806 cells. Endogenous TNFAIP2 protein was immunoprecipitated using ananti-IQGAP1 antibody. Immunoglobulin (Ig)G served as the negative control. Endogenous TNFAIP2 was detected by WB.

(C-G) IQGAP1 knockdown in HCC1806 and HCC1937 cells significantly decreased cell viability in the presence of EPI (0-1600 nM) and BMN (0-40 μM), as measured by the SRB assay. IQGAP1 protein expression was detected by WB.

(H) IQGAP1 knockdown in HCC1806 and HCC1937 cells increased DNA damage of EPI and BMN. HCC1806 and HCC1937 cells with IQGAP1 knockdown were treated with 800 nM EPI for 24h and 10 μM BMN for24h, respectively. ITGB4, γH2AX, and PARP protein expression was detected by WB.

(I-K) IQGAP1 knockdown abolished TNFAIP2-confered resistance to EPI and BMN. HCC1806 cells with stable TNFAIP2 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (0-1600 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. IQGAP1 protein expression was detected by WB.

(L) IQGAP1 knockdown abolished TNFAIP2-confered resistance to EPI and BMN. HCC1806 and HCC1937 cells with stable TNFAIP2 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (800 nM) and BMN (10 μM) for 24h, respectively. Protein expression levels were analyzed by WB.

(M) IQGAP1 knockdown abolished TNFAIP2-confered Rac1 activation. HCC1806 and HCC1937 cells with stable TNFAIP2 overexpression were transfected with IQGAP1 or control siRNA. GTP-Rac1 levels were assessed using PAK-PBD beads.

ITGB4 interacts with TNFAIP2 and promotes TNBC drug resistance and DNA damage repair

In addition to IQGAP1, TNFAIP2 may interact with ITGB4 (Fig. 4A). To validate whether TNFAIP2 interacts with ITGB4, we immunoprecipitated exogenous Flag-tagged TNFAIP2 proteins from HCC1806 cells by using Flag-M2 beads and detected endogenous ITGB4 proteins(Fig. 5A).To further confirm the protein‒protein interaction between endogenousTNFAIP2 and ITGB4proteins, we collected HCC1806 cell lysates and performed immunoprecipitation using an anti-TNFAIP2/ITGB4 antibody and found that endogenous TNFAIP2/ITGB4 protein interacted with endogenous ITGB4/TNFAIP2 protein (Fig. 5B-C).We further mapped the regions of TNFAIP2 and ITGB4 proteins responsible for the interaction(Fig. S5I) by generating a series of Flag-TNFAIP2/GST-fused TNFAIP2 deletion mutants and transfected them into HEK293T cells together with full-length GST-fused ITGB4/ITGB4. Then, we performed immunoprecipitation assays using Flag-M2 beads and glutathione beads. As shown in Figure S5J-K, the N-terminus (218-287 aa) of the TNFAIP2 (TNFAIP2-S-N1-3) protein interacted with ITGB4. To map the domains of ITGB4 that interact with TNFAIP2, we transfected Flag-tagged full-length TNFAIP2 into HEK293T cells with full-length or truncated ITGB4. We found that the C-terminus (710-740 aa) of the ITGB4 protein interacted with TNFAIP2 (Fig. S5L-M). Taken together, these results suggest that TNFAIP2 interacts with ITGB4 and that their interaction is mediated through the N-terminus of TNFAIP2 and the C-terminus of ITGB4.

ITGB4 interacts with TNFAIP2 and promotes TNBC drug resistance and DNA damage repair

(A) TNFAIP2 interacts with ITGB4.HCC1806 cells with stable TNFAIP2 overexpression were collected from flag-tagged TNFAIP2 cell lysates for immunoprecipitation assays using Flag-M2 beads, and ITGB4 was detected by WB.

(B) Endogenous TNFAIP2 interacts with ITGB4 in HCC1806 cells. Endogenous TNFAIP2 protein was immunoprecipitated using an anti-TNFAIP2 antibody. IgG served as the negative control. Endogenous ITGB4 was detected by WB.

(C) Endogenous ITGB4 interacts with TNFAIP2 in HCC1806 cells. Endogenous ITGB4 protein was immunoprecipitated using an anti-ITGB4 antibody. IgG served as the negative control. Endogenous TNFAIP2 was detected by WB.

(D-I) ITGB4 knockdown in HCC1806 and HCC1937 cells significantly decreased cell viability in the presence of EPI (0-800 nM) and BMN (0-40 μM), as measured by the SRB assay. ITGB4 protein expression was detected by WB.

(J-N) ITGB4 depletion promotes HCC1806 breast cancer cell sensitivity to EPI and BMN treatment in vivo. HCC1806 cells with stable ITGB4 knockdown were transplanted into the fat pad of 7-week-old female nude mice. When the average tumor size reached approximately 50 mm3 after inoculation, the mice in each group were randomly divided into two subgroups (n=4/group) to receive EPI (2.5 mg/kg), BMN (1 mg/kg) or vehicle control for 22 days(J). Tumor masses were collected and weighed at the end of the experiments(K), and tumor size was measured twice a week (L-N).*: P<0.05, **: P<0.01, t-test.

(O) ITGB4 knockdown increased DNA damage of EPI and BMN. HCC1806 and HCC1937 cells with ITGB4 knockdown were treated with 400 nM EPI for 24h and 5 μM BMN for 24h, respectively. ITGB4, γH2AX, and PARP protein expression was detected by WB.

To explore the function of ITGB4 in TNBC drug resistance, we knocked down ITGB4 in HCC1806 and HCC1937cells. As shown in Figure 5D-I, knockdown of ITGB4 significantly decreased cell viability in the presence of EPI and BMN in both cell lines. Knockdown of ITGB4 also suppressed HCC1806 xenograft growth in vivo. More importantly, ITGB4 knockdown further decreased tumor volume when mice were treated with EPI and BMN (Figure. 5J-N). Meanwhile, BMN treatment had no effect on the body weight of mice, but EPI treatment decreased mouse body weight due to its toxicity (Fig. S5H). We then examined the effects of ITGB4 on DNA damage repair and found that ITGB4 promotes DNA damage repair in response to EPI and BMN. ITGB4 knockdown increased γH2AX and cleaved-PARP protein expression levels when HCC1806 and HCC1937 cells were treated with EPI and BMN (Fig. 5O). Furthermore, the function of ITGB4 was validated by using two other drugs, DDP and AZD (Fig. S5A-G). These results suggested that ITGB4 increases TNBC drug resistance and promotes DNA damage repair.

ITGB4 activates Rac1 through TNFAIP2 and IQGAP1

It is well known that ITGB4 can activate Rac127 and that TNFAIP2 interacts with Rac1 and activates it9. To test whether ITGB4 activates Rac1 through TNFAIP2, we measured the levels of GTP-bound Rac1 in ITGB4-overexpressing and ITGB4-knockdown cells. Overexpression of ITGB4 significantly increased the levels of GTP-bound Rac1 in bothHCC1806 and HCC1937 cells (Fig. 6A). In agreement with this observation, knockdown of ITGB4 significantly decreased the levels of GTP-bound Rac1 in both cell lines (Fig. 6B). Next, we knocked down TNFAIP2 in ITGB4-overexpressing HCC1806 and HCC1937 cells and found that ITGB4-increased Rac1 activity was blocked by TNFAIP2 knockdown (Fig. 6C-D). Collectively, these results demonstrate that ITGB4 activates Rac1 through TNFAIP2.

ITGB4 activates Rac1 through TNFAIP2 and IQGAP1

(A) Overexpression of ITGB4 increased GTP-Rac1 levels in HCC1806 and HCC1937 cells. GTP-Rac1 were assessed using PAK-PBD beads.

(B) Knockdown of ITGB4 by shRNA decreased GTP-Rac1 levels in HCC1806 and HCC1937 cells.

(C-D) ITGB4 activates Rac1 through TNFAIP2. HCC1806 (C) and HCC1937 (D) cells with stable ITGB4 overexpression were transfected with TNFAIP2 or control siRNA.

(E-F) ITGB4 activates Rac1 through IQGAP1.HCC1806 (E) and HCC1937 (F) cells with stable ITGB4 overexpression were transfected with IQGAP1 or control siRNA.

It has been reported that Rac1 activity is promoted by IQGAP134 and that TNFAIP2 activates Rac1 through IQGAP1 (Fig. 4P). We wondered whether ITGB4 activates Rac1 through IQGAP1; therefore, we knocked down IQGAP1 in HCC1806 and HCC1937 cells with stable overexpression of ITGB4 and found that the ITGB4-induced increase in Rac1 activity was abolished by IQGAP1 knockdown (Fig. 6E-F). These results suggest thatITGB4 activates Rac1 through TNFAIP2 and IQGAP1.

ITGB4 promotes TNBC drug resistance via TNFAIP2/IQGAP1/Rac1

Since ITGB4, TNFAIP2, and IQGAP1 promote drug resistance by promoting DNA damage repair in TNBC, we wondered whether ITGB4 promoted drug resistance through the TNFAIP2/IQGAP1/Rac1 axis. We knocked down TNFAIP2, IQGAP1, and Rac1in ITGB4-overexpressing cells and found that blocking the TNFAIP2/IQGAP1/Rac1 axis increased the sensitivity of ITGB4-overexpressing HCC1806 (Fig. 7A-I) and HCC1937 cells to EPI and BMN (Fig. S6O-W).We also found that γH2AX and cleaved-PARP levels were upregulated in TNFAIP2/IQGAP1/Rac1 knockdown HCC1806 and HCC1937 cells stably expressing ITGB4 in the presence of EPI and BMN (Fig. 7J-L, Fig. S6X-Z). DDP and AZD treatment generated similar results (Fig. S6A-N). Together, these results suggest that ITGB4 promotes DNA damage repair and drug resistance via the TNFAIP2/IQGAP1/Rac1 axis.

ITGB4 promotes TNBC drug resistance via TNFAIP2/IQGAP1/Rac1

(A-C) ITGB4 promotes TNBC drug resistance through TNFAIP2. TNFAIP2 knockdown abolished ITGB4-induced resistance to EPI and BMN. HCC1806 cells with stable ITGB4 overexpression were transfected with TNFAIP2 or control siRNA, followed by treatment with EPI (0-400 nM) and BMN (0-30 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Protein expression levels were analyzed by WB.

(D-F) ITGB4 promotes TNBC drug resistance through IQGAP1. HCC1806 cells with stable ITGB4 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (0-800 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Protein expression levels were analyzed by WB.

(G-I) ITGB4 promotes TNBC drug resistance through Rac1. HCC1806 cells with stable ITGB4 overexpression were transfected with Rac1 or control siRNA, followed by treatment with EPI (0-400 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Protein expression levels were analyzed by WB.

(J) ITGB4 promotes DNA damage repair through TNFAIP2. HCC1806 cells with stable ITGB4 overexpression were transfected with TNFAIP2 or control siRNA, followed by treatment with EPI (400 nM) and BMN (5 μM) for 24h. Protein expression levels were analyzed by WB.

(K) ITGB4 promotes DNA damage repair through IQGAP1. HCC1806 cells with stable ITGB4 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (400 nM) and BMN (5 μM) for 24h. Protein expression levels were analyzed by WB.

(L) ITGB4 promotes DNA damage repair through Rac1. HCC1806 cells with stable ITGB4 overexpression were transfected with Rac1 or control siRNA, followed by treatment with EPI (400 nM) and BMN (5 μM) for 24h. Protein expression levels were analyzed by WB.

TNFAIP2 expression levels positively correlated with ITGB4 in TNBC tissues

To test whether ITGB4 and TNFAIP2 are coexpressed in TNBC, we collected 135 TNBC specimens for IHC (the IQGAP1 antibody did not work for IHC). Clinical pathological parameters, including patient age, tumor size, lymph node status at the time of diagnosis, and follow-up status, including adjuvant treatment and tumor recurrence, were retrospectively obtained from the Department of Pathology, Henan Provincial People’s Hospital, Zhengzhou University, China. We performed IHC analyses on two breast cancer tissue chips containing a total of 135 patients with TNBC (Fig. 8A-D).TNFAIP2 and ITGB4 protein expression levels were significantly positively correlated(Table 1).

TNFAIP2 expression levels positively correlated with ITGB4 in TNBC tissues

Representative IHC images of TNFAIP2 and ITGB4 protein expression in breast cancer tissues are shown. The final expression assessment was performed by combining the two scores (0–2=low, 6–7=high). A and B indicate low scores, C and D indicate high scores, and Figure E is the work model of this paper.

Discussion

Chemotherapies, including EPI and DDP, are the main choice for TNBC patients. Unfortunately, TNBC frequently develops resistance to chemotherapy39. Currently, PARP inhibitors are effective for TNBC with BRCA1/2 mutation or homologous recombination deficiency (HRD)40–42. PARP inhibitors can cause DNA damage repair defects and have synergistic lethal effects with HRD. Meanwhile, chemotherapy and PARP inhibitor resistance is also a major problem in the clinic.

In this study, we first found that TNFAIP2 promotes TNBC drug resistance and DNA damage repair through Rac1. Next, we found that TNFAIP2 interacts with IQGAP1 and ITGB4. We verified that ITGB4 promotes TNBC drug resistance and DNA damage repair through the TNFAIP2/IQGAP1/Rac1 axis.

Interestingly, we discovered for the first time that ITGB4 and TNFAIP2 promote Rac1 activity through IQGAP1. Our study reveals that ITGB4 promotes TNBC resistance through TNFAIP2-, IQGAP1-, and Rac1-mediated DNA damage repair (Figure 7). This study provides new targets for reversing TNBC resistance.

ITGB4 is well known to promote breast cancer stemness and can be activated by laminin 543. In addition, ITGB4 is generally in partner with ITGA6, which is another marker of breast cancer stem cells44 and drug resistance43. Therefore, whether ITGA6 has similar functions needs further study. It was reported that ITGB4 activates Rac145, but the mechanism is unclear. For the first time, we revealed that ITGB4 activates Rac1 through TNFAIP2 and IQGAP1. More importantly, ITGB4 promotes drug resistance through the TNFAIP2/IQGAP1/Rac1 axis.

TNFAIP2 plays important roles in different cellular and physiological processes, including cell proliferation, adhesion, migration, membrane TNT formation, angiogenesis, inflammation and tumorigenesis14. We previously found that TNFAIP2 was regulated by KLF5 and interacted with the small GTPases Rac1 and Cdc42, thereby regulating the actin cytoskeleton and cell morphology in breast cancer cells9. However, the detailed mechanism is not clearly understood. In this study, we found that IQGAP1 mediates this process.IQGAP1 is a crucial regulator of cancer development by scaffolding and facilitating different oncogenic pathways, especially Rac1/Cdc42, thus affecting proliferation, adhesion, migration, invasion, and metastasis46. In addition, IQGAP1 is increased during the differentiation of ovarian cancer stem cells and promotes aggressive behaviors47. In our study, we found that TNFAIP2 interacts with IQGAP1 and thus activates Rac1 to induce chemotherapy and PARP inhibitor drug resistance.

Furthermore, TNFAIP2 was reported to induce epithelial-to-mesenchymal transition and confer platinum resistance in urothelial cancer cells12, and in embryonic stem cell (ESC) differentiation, TNFAIP2 was found to be important in controlling lipid metabolism, which supports the ESC differentiation process and planarian organ maintenance48. Another study found that TNFAIP2 overexpression enhanced TNT-mediated autophagosome and lysosome exchange, preventing advanced glycation end product (AGE)-induced autophagy and lysosome dysfunction and apoptosis49. In cancer treatment, TNFAIP2 was chosen as one of the six signature genes predicting chemotherapeutic and immunotherapeutic efficacies, with high-senescore patients benefiting from immunotherapy and low-senescore patients responsive to chemotherapy50.

These reports provide apossible explanation for previous studies showing that ITGB4 is important in EMT and cancer stemness. According to our results that there isan interaction between ITGB4 and TNFAIP2, ITGB4 might regulate EMT and stemness through TNFAIP2.TNFAIP2 is one of the important factors induced by tumor necrosis factor alpha (TNFα). Interestingly, TNFα release could be induced by therapeutic drugs from multiple tumor cell lines. The acquisition of docetaxel resistance was accompanied by increased constitutive production of TNFα51. In addition, TNFα is a key tumor-promoting effector molecule secreted by tumor-associated macrophages. In vitro neutralizing TNFα was observed to inhibit tumor progression and improve the curative effect of bevacizumab52. Therefore, the mechanism by which TNFα plays a role in chemotherapeutic resistance in breast cancer should be further investigated. TNFα neutralizing provides novel opinions elucidating drug resistance.

For future studies, it will be important to develop Tnfaip2 knockout mice to investigate the exact role of TNFAIP2 physiologically. According to recent studies and our findings, agents targeting the interaction among ITGB4/TNFAIP2/IQGAP1 would be a promising trend for developing drugs to overcome the resistance phenomenon.

In summary, ITGB4 and TNFAIP2 play important roles in breast cancer chemoresistance. TNFAIP2 activates Rac1 to promote chemoresistance through IQGAP1. In addition, ITGB4 activates Rac1 through TNFAIP2 and IQGAP1 and confer DNA damage drug resistance in TNBC (Figure 8 E). and HCC1937 cells were cultured in RPMIThese results indicate that the ITGB4/TNFAIP2/IQGAP1/Rac1 axis provides potential therapeutic targets to overcome DNA damage drug resistance in TNBC.

Data availability statement

The authors confirm that the data supporting the findings of this study are available within the article and its supplementary materials.

eLife Assessment

This study presents a valuable finding that TNFAIP2 interacts with IQGAP1 and Integrin β4. Moreover, Integrin β4 activates Rac1 through TNFAIP2 and IQGAP1 and confers drug resistance in triple-negative breast cancer. The evidence supporting the claims of the authors is solid. The work will be of interest to medical biologists working on breast cancer or drug resistance.

Reviewer #1 (Public Review):

In this study, Fang H et al. describe a potential pathway, ITGB4-TNFAIP2-IQGAP1-Rac1, that may involve in the drug resistance in triple negative breast cancer (TNBC). Mechanistically, it was demonstrated that TNFAIP2 bind with IQGAP1 and ITGB4 activating Rac1 and the following drug resistance. The present study focused on breast cancer cell lines with supporting data from mouse model and patient breast cancer tissues. The study is interesting. The experiments were well controlled and carefully carried out. The conclusion is supported by strong evidence provided in the manuscript. The authors may want to discuss the link between ITGB4 and Rac 1, between IQGAP1 and Rac1, and between TNFAIP2 and Rac1 as compared with the current results obtained. This is important considering some recent publications in this area (Cancer Sci 2021, J Biol Chem 2008, Cancer Res 2023).

Reviewer #2 (Public Review):

Breast cancer is the most common malignant tumor in women. One of subtypes in breast cancer is so called triple-negative breast cancer (TNBC), which represents the most difficult subtype to treat and cure in the clinic. Chemotherapy drugs including epirubicin and cisplatin are widely used for TNBC treatment. However, drug resistance remains as a challenge in the clinic. The authors uncovered a molecular pathway involved in chemotherapy drug resistance, and molecular players in this pathway represent as potential drug targets to overcome drug resistance. The experiments are well designed and the conclusions drawn mostly were supported by the data. The findings have potential to be translated into the clinic.

Reviewer #3 (Public Review):

In this manuscript, Fang and colleagues found that IQGAP1 interacts with TNFAIP2, which activates Rac1 to promote drug resistance in TNBC. Furthermore, they found that ITGB4 could interact with TNFAIP2 to promote TNBC drug resistance via the TNFAIP2/IQGAP1/Rac1 axis by promoting DNA damage repair.

This work has good innovation and high potential clinical significance.

Author Response

The following is the authors’ response to the original reviews.

Reviewer #1:

In this study, Fang H et al. describe a potential pathway, ITGB4-TNFAIP2-IQGAP1-Rac1, that may involve in the drug resistance in triple negative breast cancer (TNBC). Mechanistically, it was demonstrated that TNFAIP2 bind with IQGAP1 and ITGB4 activating Rac1 and the following drug resistance. The present study focused on breast cancer cell lines with supporting data from mouse model and patient breast cancer tissues. The study is interesting. The experiments were well controlled and carefully carried out. The conclusion is supported by strong evidence provided in the manuscript. The authors may want to discuss the link between ITGB4 and Rac1, between IQGAP1 and Rac1, and between TNFAIP2 and Rac1 as compared with the current results obtained. This is important considering some recent publications in this area (Cancer Sci 2021, J Biol Chem 2008, Cancer Res 2023). In addition, some key points need to be addressed in order to support their conclusion in full.

Thanks for your positive comments.

1. It is rarely found studies using the term of "DNA damage drug resistance". Do the authors mean "DNA damage and drug resistance" or "DNA damage-related drug resistance" or "DNA damage-induced drug resistance"? It is better to define "DNA damage drug resistance" in the manuscript if it is not a common term in the field.

We agree with you that the description "DNA damage-related drug resistance" is better so that we revised it uniformly in the manuscript.

1. For Figure 4A, it is stated the IQGAP1 is identified via IP-MS. However, the MS results are not presented in the Figure or in the supplementary. In Figure 4A, only the IP results with silver staining was presented. Moreover, based on the silver staining here, a bunch of proteins were increased in TNFAIP2 overexpression group compared to the vector group. Especially, there is a much clearer band at 52kDa. The authors didn't explain why they chose IQGAP1 and ITGB4 which are less clear than the protein(s) at 52kDa.

Supplementary table 1 is our mass spectrometry results. There are two reasons for choosing ITGB4 and IQGAP1. Firstly, we selected the proteins that indeed interact with TNFAIP2 according to our verification experiments. Secondly, we were interested in the mechanism by which TNFAIP2 promoting DNA damage-related drug resistance, and we found that ITGB4 promoted drug resistance, while IQGAP1 activated Rac1.

1. According to the images in Figure 4C, the efficiency of si-IQGAP1 is limited. The authors could analyze the WB image to confirm the inhibition efficiency of si-IQGAP1.

We analyzed the WB images and the quantitative results are as follows in Author response image 1. The knockdown efficiency is acceptable.

Author response image 1.

1. In Figure 5B, I wonder whether the authors can explain why the IgG could immunoprecipitate similar amount of ITGB4 protein as input group.

In this experiment, the Input group had relatively less loading amount (5%), while the IgG group had nonspecific binding.

1. According to the results from Figure 6B, the inhibition efficiency of shITGB4#1 is much higher than shITGB4#2. However, the effects of shITGB4#1 on GTP-Rac1 are similar to or even weaker than those of shITGB4#2 in both HCC1806 and HCC1937. Can this be explained?

The possible reason is that downregulation of ITGB4 expression to a certain level is sufficient to inhibit the activation of Rac1.

1. In Figure 6F, there are double bands for ITGB4 while only one band shows in other Figures. Please find a better representative image here.

ITGB4 has a cleaved band in addition to the main band. These two bands could be separated when we used a low concentration SDS-PAGE gel.

1. In the manuscript, GAPDH, b-Actin and Tubulin are used in different experiments as internal controls. Is there any specific reason to using different internal controls for different experiments here?

There is no specific reason using different internal controls. These experiments were conducted by different person. Each individual chose different internal controls based on the protein sizes.

1. I cannot find Table 1 for the correlation results for TNFAIP2 and ITGB4. I wonder whether Figure 8E is the Table 1 as is mentioned, since it is stated in line 561 that Figure 8E is "the work model of this paper" but actually Figure 8F is. If Figure 8E is the correlation results, I highly recommended the scatter plots graph is used here to present more clear and visualized correlation between TNFAIP2 and ITGB4.

Figure 8E is indeed the correlation result. In addition, Figure 8E could not be presented as scatter plot graph because the pattern of TNFAIP2 and ITGB4 expression is negative or positive according to the determination of IHC results which was carried out by professional pathologists.

1. Throughout the whole manuscript, no description of N number was found in figure legends or in Methods for in vitro experiments. N number is important for statistical analysis.

All our experiments have set up three replicates. We provide this information in figure legends.

Reviewer #2:

Breast cancer is the most common malignant tumor in women. One of subtypes in breast cancer is so called triple-negative breast cancer (TNBC), which represents the most difficult subtype to treat and cure in the clinic. Chemotherapy drugs including epirubicin and cisplatin are widely used for TNBC treatment. However, drug resistance remains as a challenge in the clinic. The authors uncovered a molecular pathway involved in chemotherapy drug resistance, and molecular players in this pathway represent as potential drug targets to overcome drug resistance. The experiments are well designed and the conclusions drawn mostly were supported by the data. The findings have potential to be translated into the clinic.

Thanks for your positive comments.

1. In Introduction, the statement of "Breast cancer is the most common malignant tumor in women, and the morbidity and mortality rates of female malignant tumors are ranked first in the world" is inaccurate.

We have revised the description as“Breast cancer is the most commonly diagnosed cancer and the leading cause of cancer death in women”.

1. In Materials and Methods, "Immunopurification and silver staining" is not correct, which should be replaced with "Immunoprecipitation and silver staining".

We replaced the description in the manuscript according to your suggestion.

1. It is unclear Why the authors chose the two TNBC cell lines, HCC1806 and HCC1937, for cell models in this work.

We chose these two cell lines according to our previous work“KLF5 promotes breast cancer proliferation, migration and invasion in part by upregulating the transcription of TNFAIP2” (doi: 10.1038/onc.2015.263. Epub 2015 Jul 20).

1. To demonstrate TNFAIP2 and ITGB4 confer TNBC drug resistance in vivo, the knockdown efficiency of animal experiments was not shown.

The knockdown efficiency of animal experiments was shown below. We added this result into Figure 2-figure supplement 2G and Figure 5-figure supplement 2N.

1. I would strongly suggest the authors seek help from a language editing service to improve the manuscript.

We improved the manuscript by using a professional English language editing service and we have carefully revised the manuscript.

Reviewer #3:

In this manuscript, Fang and colleagues found that IQGAP1 interacts with TNFAIP2, which activates Rac1 to promote drug resistance in TNBC. Furthermore, they found that ITGB4 could interact with TNFAIP2 to promote TNBC drug resistance via the TNFAIP2/IQGAP1/Rac1 axis by promoting DNA damage repair.

This work has good innovation and high potential clinical significance. However, there are several unsolved concerns that have to be addressed.

Thanks for your positive comments.

1. In the manuscript, there are four drugs used for in vitro cell experiments, why is olaparib (AZD) not used for in vivo animal experiments？

There are two reasons why we did not choose AZD. First，the killing effect of AZD is not as strong as that of BMN. Second, AZD is more expensive than BMN. We finally chose BMN for animal experiments.

1. In Figure 4B, why the immunoprecipitation experiments is done in HCC1806 cell line？

In our previous study “KLF5 promotes breast cancer proliferation, migration and invasion in part by upregulating the transcription of TNFAIP2” (doi: 10.1038/onc.2015.263. Epub 2015 Jul 20), we found that TNFAIP2 knockdown could obviously inhibit the activation of Rac1 in HCC1806 when compared to the result in HCC1937. So, we used HCC1806 cell line to perform the IP-Mass assay.

1. There should be data showing the knockdown effect of TNFAIP2 and ITGB4 in animal experiments.

We addressed the same question above (Reviewer #2, Question#4).

1. When screening the interaction regions between ITGB4 and TNFAIP2, why the TNFAIP2 protein truncation strategy is to delete the N-terminus？

In fact, we also deleted the C-terminus, but the deletion of C-terminus of TNFAIP2 did not affect the interaction.

1. In the manuscript, "input" should be changed to "Input".

We corrected it.

1. There should be a space between "Figure" and numbers.

We add a space between "Figure" and numbers.

Version 2

Abstract

Abstract

Anti-tumor drug resistance is a challenge for human triple-negative breast cancer treatment. Our previous work demonstrated that TNFAIP2 activates RAC1 to promote triple-negative breast cancer cell proliferation and migration. However, the mechanism by which TNFAIP2 activates RAC1 is unknown. In this study, we found that TNFAIP2 interacts with IQGAP1 and Integrin β4. Integrin β4 activates RAC1 through TNFAIP2 and IQGAP1 and confers DNA damage-related drug resistance in triple-negative breast cancer. These results indicate that the Integrin β4/TNFAIP2/IQGAP1/RAC1 axis provides potential therapeutic targets to overcome DNA damage-related drug resistance in triple-negative breast cancer.

Introduction

Breast cancer is the most commonly diagnosed cancer and the leading cause of cancer death in women[1]. Although the diagnosis and treatment of breast cancer has entered the era of molecular typing, 35% of breast cancers still experience recurrence, metastasis and treatment failure[2]. According to the expression of estrogen receptor (ERα), progesterone receptor (PR) and human epidermal growth factor receptor (HER2), breast cancer is divided into ER/PR-positive, HER2-positive and triple-negative breast cancer (TNBC)[3]. For ER/PR-positive and HER2-positive breast cancer, endocrine therapies such as tamoxifen and anti-HER2 targeted therapy such as trastuzumab have achieved good efficacy. Targeted drugs for TNBC patients with BRCA1/2 mutations include two PARP inhibitors, olaparib and talazoparib. These targeted drugs cannot fully meet the clinical needs of patients with various TNBC subtypes[4]. Currently, DNA damage chemotherapy drugs, including epirubicin and cisplatin, are widely used for TNBC treatment.

TNBC often recurs and metastasizes due to the development of chemoresistance, although it is initially responsive to chemotherapeutic drugs[5]. Chemoresistance severely impacts the clinical outcomes of patients. Tumor cells become resistant to chemotherapeutic agents through several mechanisms, such as improving DNA damage repair, changing the intracellular accumulation of drugs, or increasing anti-apoptotic mechanisms[6]. Therefore, characterization of the underlying molecular mechanisms by which resistance occurs will provide opportunities to develop precise therapies to enhance the efficacy of standard chemotherapy regimens [7, 8].

TNFAIP2 is abnormally highly expressed in a variety of tumor cells, including TNBC[9], nasopharyngeal carcinoma[10], malignant glioma[11], uroepithelial carcinoma[12] and esophageal squamous cell carcinoma[13], and is associated with poor prognosis. Our previous work [9, 14]showed that TNFAIP2, as a KLF5 downstream target protein, can interact with RAC1[15], a member of the Rho small GTP enzyme family, and activate RAC1 to alter the cytoskeleton, thereby inducing filopodia and lamellipodia formation and promoting the adhesion, migration and invasion of TNBC cells. After activation, RAC1 can activate AKT, PAKs, NADPH oxidase and other related signaling pathways to promote cell survival, proliferation, adhesion, migration and invasion[16]. Activation of RAC1 can reduce the therapeutic response to trastuzumab in breast cancer and increase the resistance of TNBC cells to paclitaxel[17], but the specific mechanism of action is not completely clear.

RAC1 has been shown to play an important role in DNA damage repair. Activated RAC1 can promote the phosphorylation of the DNA damage response-related molecules ATM/ATR, CHK1/2 and H2AX by activating the activity of protein kinases such as ERK1/2, JNK and p38 [18, 19], thus improving the DNA damage repair ability and inhibiting tumor cell apoptosis[20–22]. RAC1 also promotes aldolase release and activation by changing the cytoskeleton and activates the ERK pathway to increase the pentose phosphate pathway to promote nucleic acid synthesis, providing more raw materials for DNA damage repair[23, 24]. At the same time, the interaction of RAC1 and PI3K promoted AKT phosphorylation and glucose uptake[25, 26]. Therefore, RAC1 is well established to promote the chemoresistance of breast cancer by promoting DNA damage repair.

Integrin β4 (ITGB4) is a major component of hemidesmosomes and a receptor molecule of laminin. Studies have shown that laminin-5 interacts with ITGB4 to activate RAC1 activity and promote cell migration[27] and polarization[28] by altering the cytoskeleton. Since ITGB4-positive cancer stem cell (CSC)-enriched mesenchymal cells were found to reside in an intermediate epithelial/mesenchymal phenotypic state, ITGB4 can be used to enable stratification of mesenchymal-like TNBC cells[29]. In addition, the expression ofITGB4 on ALDHhigh breast cancer and head and neck cancer cells was significantly greater than that on ALDHlow cells, proving the effects that ITGB4 targets on both bulk and CSC populations[30]. Furthermore, ITGB4-overexpressing TNBC cells provided cancer-associated fibroblasts (CAFs) with ITGB4 proteins via exosomes, and ITGB4-overexpressing CAF-conditioned medium promoted the proliferation, epithelial-to-mesenchymal transition, and invasion of breast cancer cells[31]. ITGB4 also promotes breast cancer cell resistance to tamoxifen-induced apoptosis by activating the PI3K/AKT signaling pathway and promotes breast cancer cell resistance to anoikis by activating RAC1[32]. However, how ITGB4 activates RAC1 is not completely clear.

RAC1 activity is regulated by guanylate exchange factors (GEFs), GTPase activation proteins (GAPs), and guanine separation inhibitors (GDIs)[33]. GAPs typically provide the necessary catalytic groups for GTP hydrolysis, but not all GAPs function as hydrolases. IQGAP1 lacks an arginine in the GTPase binding domain and cannot exert the hydrolysis effect of GAPs[34]. IQGAP1 can increase the activity of RAC1 and CDC42[35, 36].

In this study, we demonstrated that TNFAIP2 interacts with IQGAP1 and ITGB4. ITGB4 promotes TNBC drug resistance via the TNFAIP2/IQGAP1/RAC1 axis by promoting DNA damage repair. Our results suggest that ITGB4 and TNFAIP2 might serve as promising therapeutic targets for TNBC.

Results

TNFAIP2 promotes TNBC DNA damage-related drug resistance

To explore the functional significance of TNFAIP2 in TNBC drug resistance, we constructed stable TNFAIP2 overexpression and TNFAIP2 knockdown HCC1806 and HCC1937 cells. As shown in Figure 1A-E, overexpression of TNFAIP2 significantly increased cell viability when treated with EPI and BMN. Additionally, knockdown of TNFAIP2 significantly decreased cell viability when treated with EPI and BMN (Figure 1F-J). We then examined the effects of TNFAIP2 on DNA damage repair and found that TNFAIP2 promotes DNA damage repair in response to EPI and BMN. TNFAIP2 overexpression decreased the protein expression levels of γH2AX, a marker of DNA damage, and cleaved-PARP, a marker of apoptosis (Figure 1K). Additionally, knockdown of TNFAIP2 significantly increased γH2AX and cleaved-PARP protein expression levels in response to EPI and BMN in both cell lines (Figure 1L).

TNFAIP2 promotes TNBC DNA damage-related drug resistance

(A-E) Stable TNFAIP2 overexpression in HCC1806 and HCC1937 cells significantly increased cell viability in the presence of EPI (0-1.6 μM) or BMN (0-40 μM) treatment for 48 h, as measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. TNFAIP2 protein expression was detected by WB.

(F-J) Stable TNFAIP2 knockdown in HCC1806 and HCC1937 cells significantly decreased cell viability in the presence of EPI (0-1.6 μM) or BMN (0-40 μM) treatment for 48 h, as measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. TNFAIP2 protein expression was detected by WB.

(K) TNFAIP2 promoted DNA damage repair in the presence of EPI and BMN. HCC1806 and HCC1937 cells stably overexpressing TNFAIP2 were treated with 400 or 800 nM EPI for 48 h and 10 μM BMN for 24 h, respectively. TNFAIP2, γH2AX, and PARP protein expression was detected by WB.

(L) TNFAIP2 knockdown increased DNA damage in the presence of EPI and BMN. Stable TNFAIP2 knockdown cells were treated with 400 or 800 nM EPI for 24 or 48h and 2.5 μM BMN for 24 h. TNFAIP2, γH2AX, and PARP protein expression was detected by WB.

Figure 1-source data 1

Uncropped western blot images for Figure 1

The function of TNFAIP2 was further validated by using two other DNA damage drugs, DDP and AZD (Figure 1-figure supplement 1A-L). These results suggested that TNFAIP2 enhances TNBC cell drug resistance by promoting DNA damage repair.

TNFAIP2 confers TNBC drug resistance in vivo

To test whether TNFAIP2 also decreases the sensitivity of TNBC cells to EPI and BMN in vivo, we performed animal experiments in nude mice. HCC1806 cells with stable TNFAIP2 knockdown were orthotopically inoculated into the fat pad of 7-week-old female mice (n=8 or 12/group). Western blotting was performed to detect the knockdown effect of TNFAIP2 protein in animal experiments (Figure 2-figure supplement 2G). When the tumor mass reached approximately 50 mm3, each group was divided into two subgroups to receive either EPI (2.5 mg/kg, twice a week) or vehicle control for 23 days and either BMN (1 mg/kg, twice a week) or vehicle control for 29 days. We observed that depletion of TNFAIP2 suppressed breast cancer cell growth in vivo. This is consistent with our previous report[9]. More importantly, TNFAIP2 depletion further decreased tumor volume when mice were treated with EPI and BMN (Figure 2A-F). Meanwhile, BMN treatment had no effect on the body weight of mice (Figure 2-figure supplement 1F). Consistently, EPI and DDP generated similar results but decreased mouse body weight due to their high toxicity (Figure 2-figure supplement 1D-E). These results suggest that inhibition of TNFAIP2 expression can overcome HCC1806 breast cancer cell drug resistance in animals.

TNFAIP2 confers TNBC drug resistance in vivo

(A-F) TNFAIP2 knockdown increased the sensitivity ofHCC1806 breast cancer cells to EPI and BMN in vivo. HCC1806 cells with stable TNFAIP2 knockdown were transplanted into the fat pad of 7-week-old female nude mice. When the average tumor size reached approximately 50 mm3 after inoculation, mice in each group were randomly divided into two subgroups (n=4/group) to receive EPI (2.5 mg/kg), BMN (1 mg/kg) or vehicle control for 23 or 29 days (A-B). Tumor size was measured twice a week (C-D), and tumor masses were collected and weighed at the end of the experiments (E-F). *: P<0.05, **: P<0.01, ***: P<0.001, t-test.

TNFAIP2 promotes TNBC drug resistance and DNA damage repair via RAC1

Since chemotherapeutic agents and PRAP inhibitors induce DNA damage directly or indirectly, DNA damage repair ability profoundly affects the sensitivity of cancer cells to these therapies[37, 38]. Since TNFAIP2 can activate RAC1, a well-known drug resistance protein, we investigated whether TNFAIP2 induces chemotherapeutic resistance through RAC1. We found that RAC1 knockdown abrogated the effects of TNFAIP2 overexpression-induced drug resistance to EPI and BMN in HCC1806 and HCC1937 cells (Figure 3A-F). We also found that γH2AX and cleaved-PARP protein levels were up-regulated again in RAC1 knockdown and TNFAIP2-overexpressing HCC1806 and HCC1937 cells in response to EPI and BMN (Figure 3G-J). We obtained similar results by using DDP and AZD treatment (Figure 3-figure supplement 1A-J). Collectively, these results suggest that TNFAIP2 promotes DNA damage repair and drug resistance via RAC1.

TNFAIP2 promotes TNBC drug resistance and DNA damage repair via RAC1

(A-F) RAC1 knockdown abolished TNFAIP2-induced TNBC resistance to EPI and BMN. HCC1806 (A-C) and HCC1937 (D-F) cells with stable TNFAIP2 overexpression were transfected with RAC1 or control siRNA, followed by treatment with EPI (0-1600 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3-9, * P<0.05, ** P<0.01, *** P<0.001. Protein expression levels were analyzed by WB.

(G-J) RAC1 depletion abolished TNFAIP2-induced DNA damage decrease in response to EPI and BMN. HCC1806 (G-H) and HCC1937 (I-J) cells with stable TNFAIP2 overexpression were transfected with RAC1 or control siRNA, followed by treatment with EPI (400 or 800 nM) and BMN (10 μM) for 24 h, respectively. Protein expression levels were analyzed by WB.

Figure 3-source data 1

Uncropped western blot images for Figure 3

IQGAP1 mediates RAC1 activation by TNFAIP2 and promotes TNBC drug resistance

To characterize the mechanism by which TNFAIP2 activates RAC1, we performed an IP-MS experiment. We found that TNFAIP2 interacts with IQGAP1 (Figure 4A). To validate whether TNFAIP2 interacts with IQGAP1, we constructed HCC1806 cells with stable Flag-TNFAIP2 overexpression and collected Flag-tagged TNFAIP2 cell lysates for immunoprecipitation assays using Flag-M2 beads (Figure 4-figure supplement 1A).We performed immunoprecipitation using ananti-IQGAP1 antibody and found that endogenous IQGAP1 protein interacted with endogenous TNFAIP2 protein inHCC1806 cells (Figure 4B). Next, we mapped the regions of TNFAIP2 and IQGAP1 proteins responsible for the interaction by generating a series of Flag-TNFAIP2 deletion mutants and transfected them into HEK293T cells together with full-length IQGAP1. Then, we performed immunoprecipitation assays using Flag-M2 beads (Figure 4-figure supplement 1B). We demonstrated that the N-terminus (1-79 aa) of the TNFAIP2 protein interacted with IQGAP1. To explore the function of IQGAP1 in TNBC drug resistance, we knocked down IQGAP1 in HCC1806 and HCC1937 cells. As shown in Figure 4C-G, knockdown of IQGAP1 significantly decreased cell viability in the presence of EPI and BMN in both cell lines. We also examined the effects of IQGAP1 on DNA damage repair and found that IQGAP1 promotes DNA damage repair. IQGAP1 knockdown increased γH2AX and cleaved-PARP protein expression levels when HCC1806 and HCC1937 cells were treated with EPI and BMN (Figure 4H). Next, we found that IQGAP1 knockdown abrogated the effects of TNFAIP2 overexpression on resistance to EPI and BMN (Figure 4I-K, Figure 4-figure supplement 1C-E). We also found that γH2AX and cleaved-PARP protein levels were up-regulated in IQGAP1 knockdown and TNFAIP2-overexpressing HCC1806 and HCC1937 cells (Figure 4L). In addition, we found that the TNFAIP2 overexpression-induced increase in RAC1 activity was abolished by IQGAP1 knockdown (Figure 4M).

IQGAP1 mediates RAC1 activation by TNFAIP2 and promotes TNBC drug resistance

(A) The IP-MS result of TNFAIP2 in HCC1806 cells.

(B) Endogenous TNFAIP2 interacts with IQGAP1 in HCC1806 cells. Endogenous TNFAIP2 protein was immunoprecipitated using ananti-IQGAP1 antibody. Immunoglobulin (Ig)G served as the negative control. Endogenous TNFAIP2 was detected by WB.

(C-G) IQGAP1 knockdown in HCC1806 and HCC1937 cells significantly decreased cell viability in the presence of EPI (0-1600 nM) and BMN (0-40 μM), as measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3-6, * P<0.05, ** P<0.01, *** P<0.001. IQGAP1 protein expression was detected by WB.

(H) IQGAP1 knockdown in HCC1806 and HCC1937 cells increased DNA damage of EPI and BMN. HCC1806 and HCC1937 cells with IQGAP1 knockdown were treated with 800 nM EPI for 24 h and 10 μM BMN for 24 h, respectively. ITGB4, γH2AX, and PARP protein expression was detected by WB.

(I-K) IQGAP1 knockdown abolished TNFAIP2-confered resistance to EPI and BMN. HCC1806 cells with stable TNFAIP2 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (0-1600 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. IQGAP1 protein expression was detected by WB.

(L) IQGAP1 knockdown abolished TNFAIP2-confered resistance to EPI and BMN. HCC1806 and HCC1937 cells with stable TNFAIP2 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (800 nM) and BMN (10 μM) for 24 h, respectively. Protein expression levels were analyzed by WB.

(M) IQGAP1 knockdown abolished TNFAIP2-confered RAC1 activation. HCC1806 and HCC1937 cells with stable TNFAIP2 overexpression were transfected with IQGAP1 or control siRNA. GTP-RAC1 levels were assessed using PAK-PBD beads.

Figure 4-source data 1

Uncropped western blot images for Figure 4

ITGB4 interacts with TNFAIP2 and promotes TNBC drug resistance and DNA damage repair

In addition to IQGAP1, TNFAIP2 may interact with ITGB4 (Figure 4A). To validate whether TNFAIP2 interacts with ITGB4, we immunoprecipitated exogenous Flag-tagged TNFAIP2 proteins from HCC1806 cells by using Flag-M2 beads and detected endogenous ITGB4 proteins (Figure 5A).To further confirm the protein protein interaction between endogenous TNFAIP2 and ITGB4 proteins, we collected HCC1806 cell lysates and performed immunoprecipitation using an anti-TNFAIP2/ITGB4 antibody and found that endogenous TNFAIP2/ITGB4 protein interacted with endogenous ITGB4/TNFAIP2 protein (Figure 5B-C). We further mapped the regions of TNFAIP2 and ITGB4 proteins responsible for the interaction (Figure 5-figure supplement 2I) by generating a series of Flag-TNFAIP2/GST-fused TNFAIP2 deletion mutants and transfected them into HEK293T cells together with full-length GST-fused ITGB4/ITGB4. Then, we performed immunoprecipitation assays using Flag-M2 beads and glutathione beads. As shown in Figure 5-figure supplement 2J-K, the N-terminus (218-287aa) of the TNFAIP2 (TNFAIP2-S-N1-3) protein interacted with ITGB4. To map the domains of ITGB4 that interact with TNFAIP2, we transfected Flag-tagged full-length TNFAIP2 into HEK293T cells with full-length or truncated ITGB4. We found that the C-terminus (710-740 aa) of the ITGB4 protein interacted with TNFAIP2 (Figure 5-figure supplement 2L-M). Taken together, these results suggest that TNFAIP2 interacts with ITGB4 and that their interaction is mediated through the N-terminus of TNFAIP2 and the C-terminus of ITGB4.

ITGB4 interacts with TNFAIP2 and promotes TNBC drug resistance and DNA damage repair

(A) TNFAIP2 interacts with ITGB4.HCC1806 cells with stable TNFAIP2 overexpression were collected from flag-tagged TNFAIP2 cell lysates for immunoprecipitation assays using Flag-M2 beads, and ITGB4 was detected by WB.

(B) Endogenous TNFAIP2 interacts with ITGB4 in HCC1806 cells. Endogenous TNFAIP2 protein was immunoprecipitated using an anti-TNFAIP2 antibody. IgG served as the negative control. Endogenous ITGB4 was detected by WB.

(C) Endogenous ITGB4 interacts with TNFAIP2 in HCC1806 cells. Endogenous ITGB4 protein was immunoprecipitated using an anti-ITGB4 antibody. IgG served as the negative control. Endogenous TNFAIP2 was detected by WB.

(D-I) ITGB4 knockdown in HCC1806 and HCC1937 cells significantly decreased cell viability in the presence of EPI (0-800 nM) and BMN (0-40 μM), as measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. ITGB4 protein expression was detected by WB.

(J-N) ITGB4 depletion promotes HCC1806 breast cancer cell sensitivity to EPI and BMN treatment in vivo. HCC1806 cells with stable ITGB4 knockdown were transplanted into the fat pad of 7-week-old female nude mice. When the average tumor size reached approximately 50 mm3 after inoculation, the mice in each group were randomly divided into two subgroups (n=4/group) to receive EPI (2.5 mg/kg), BMN (1 mg/kg) or vehicle control for 22 days (J). Tumor masses were collected and weighed at the end of the experiments (K), and tumor size was measured twice a week (L-N).*: P<0.05, **: P<0.01, ***: P<0.001, t-test.

(O) ITGB4 knockdown increased DNA damage of EPI and BMN. HCC1806 and HCC1937 cells with ITGB4 knockdown were treated with 400 nM EPI for 24 h and 5 μM BMN for 24 h, respectively. ITGB4, γH2AX, and PARP protein expression was detected by WB.

Figure 5-source data 1

Uncropped western blot images for Figure 5

To explore the function of ITGB4 in TNBC drug resistance, we knocked down ITGB4 in HCC1806 and HCC1937 cells. As shown in Figure 5D-I, knockdown of ITGB4 significantly decreased cell viability in the presence of EPI and BMN in both cell lines. Knockdown of ITGB4 also suppressed HCC1806 xenograft growth in vivo. The knockdown effect of ITGB4 protein in animal experiments was confirmed by Western blotting (Figure 5-figure supplement 2N). More importantly, ITGB4 knockdown further decreased tumor volume when mice were treated with EPI and BMN (Figure 5J-N). Meanwhile, BMN treatment had no effect on the body weight of mice, but EPI treatment decreased mouse body weight due to its toxicity (Figure 5-figure supplement 1H). We then examined the effects of ITGB4 on DNA damage repair and found that ITGB4 promotes DNA damage repair in response to EPI and BMN. ITGB4 knockdown increased γH2AX and cleaved-PARP protein expression levels when HCC1806 and HCC1937 cells were treated with EPI and BMN (Figure 5O). Furthermore, the function of ITGB4 was validated by using two other drugs, DDP and AZD (Figure 5-figure supplement 1A-G). These results suggested that ITGB4 increases TNBC drug resistance and promotes DNA damage repair.

ITGB4 activates RAC1 through TNFAIP2 and IQGAP1

It is well known that ITGB4 can activate RAC1[27] and that TNFAIP2 interacts with RAC1 and activates it[9]. To test whether ITGB4 activates RAC1 through TNFAIP2, we measured the levels of GTP-bound RAC1 in ITGB4-overexpressing and ITGB4-knockdown cells. Overexpression of ITGB4 significantly increased the levels of GTP-bound RAC1 in both HCC1806 and HCC1937 cells (Figure 6A). In agreement with this observation, knockdown of ITGB4 significantly decreased the levels of GTP-bound RAC1 in both cell lines (Figure 6B). Next, we knocked down TNFAIP2 in ITGB4-overexpressing HCC1806 and HCC1937 cells and found that ITGB4-increased RAC1 activity was blocked by TNFAIP2 knockdown (Figure 6C-D). Collectively, these results demonstrate that ITGB4 activates RAC1 through TNFAIP2.

ITGB4 activates RAC1 through TNFAIP2 and IQGAP1

(A) Overexpression of ITGB4 increased GTP-RAC1 levels in HCC1806 and HCC1937 cells. GTP-RAC1 were assessed using PAK-PBD beads.

(B) Knockdown of ITGB4 by shRNA decreased GTP-RAC1 levels in HCC1806 and HCC1937 cells.

(C-D) ITGB4 activates RAC1 through TNFAIP2. HCC1806 (C) and HCC1937

(D) cells with stable ITGB4 overexpression were transfected with TNFAIP2 or control siRNA.

(E-F) ITGB4 activates RAC1 through IQGAP1. HCC1806 (E) and HCC1937 (F) cells with stable ITGB4 overexpression were transfected with IQGAP1 or control siRNA.

Figure 6-source data 1

Uncropped western blot images for Figure 6

It has been reported that RAC1 activity is promoted by IQGAP1[34] and that TNFAIP2 activates RAC1 through IQGAP1 (Figure 4P). We wondered whether ITGB4 activates RAC1 through IQGAP1; therefore, we knocked down IQGAP1 in HCC1806 and HCC1937 cells with stable overexpression of ITGB4 and found that the ITGB4-induced increase in RAC1 activity was abolished by IQGAP1 knockdown (Figure 6E-F). These results suggest that ITGB4 activates RAC1 through TNFAIP2 and IQGAP1.

ITGB4 promotes TNBC drug resistance via TNFAIP2/IQGAP1/RAC1

Since ITGB4, TNFAIP2, and IQGAP1 promote drug resistance by promoting DNA damage repair in TNBC, we wondered whether ITGB4 promoted drug resistance through the TNFAIP2/IQGAP1/RAC1 axis. We knocked down TNFAIP2, IQGAP1, and RAC1 in ITGB4-overexpressing cells and found that blocking the TNFAIP2/IQGAP1/RAC1 axis increased the sensitivity of ITGB4-overexpressing HCC1806 (Figure 7A-I) and HCC1937 cells to EPI and BMN (Figure 7-figure supplement 2O-W). We also found that γH2AX and cleaved-PARP levels were upregulated in TNFAIP2/IQGAP1/RAC1 knockdown HCC1806 and HCC1937 cells stably expressing ITGB4 in the presence of EPI and BMN (Figure 7J-L, Figure 7-figure supplement 2X-Z). DDP and AZD treatment generated similar results (Figure 7-figure supplement 1A-N). Together, these results suggest that ITGB4 promotes DNA damage repair and drug resistance via the TNFAIP2/IQGAP1/RAC1 axis.

ITGB4 promotes TNBC drug resistance via TNFAIP2/IQGAP1/RAC1

(A-C) ITGB4 promotes TNBC drug resistance through TNFAIP2. TNFAIP2 knockdown abolished ITGB4-induced resistance to EPI and BMN. HCC1806 cells with stable ITGB4 overexpression were transfected with TNFAIP2 or control siRNA, followed by treatment with EPI (0-400 nM) and BMN (0-30 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. Protein expression levels were analyzed by WB.

(D-F) ITGB4 promotes TNBC drug resistance through IQGAP1. HCC1806 cells with stable ITGB4 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (0-800 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. Protein expression levels were analyzed by WB.

(G-I) ITGB4 promotes TNBC drug resistance through RAC1. HCC1806 cells with stable ITGB4 overexpression were transfected with RAC1 or control siRNA, followed by treatment with EPI (0-400 nM) and BMN (0-40 μM) for 48 or 72 h, respectively. Cell viability was measured by the SRB assay. Statistical analysis was performed using one-way ANOVA, n=3, * P<0.05, ** P<0.01, *** P<0.001. Protein expression levels were analyzed by WB.

(J) ITGB4 promotes DNA damage repair through TNFAIP2. HCC1806 cells with stable ITGB4 overexpression were transfected with TNFAIP2 or control siRNA, followed by treatment with EPI (400 nM) and BMN (5 μM) for 24 h. Protein expression levels were analyzed by WB.

(K) ITGB4 promotes DNA damage repair through IQGAP1. HCC1806 cells with stable ITGB4 overexpression were transfected with IQGAP1 or control siRNA, followed by treatment with EPI (400 nM) and BMN (5 μM) for 24 h. Protein expression levels were analyzed by WB.

(L) ITGB4 promotes DNA damage repair through RAC1. HCC1806 cells with stable ITGB4 overexpression were transfected with RAC1 or control siRNA, followed by treatment with EPI (400 nM) and BMN (5 μM) for 24 h. Protein expression levels were analyzed by WB.

Figure 7-source data 1

Uncropped western blot images for Figure 7

TNFAIP2 expression levels positively correlated with ITGB4 in TNBC tissues

To test whether ITGB4 and TNFAIP2 are co-expressed in TNBC, we collected 135 TNBC specimens for IHC (the IQGAP1 antibody did not work for IHC). Clinical pathological parameters, including patient age, tumor size, lymph node status at the time of diagnosis, and follow-up status, including adjuvant treatment and tumor recurrence, were retrospectively obtained from the Department of Pathology, Henan Provincial People’s Hospital, Zhengzhou University, China. We performed IHC analyses on two breast cancer tissue chips containing a total of 135 patients with TNBC (Figure 8A-D). TNFAIP2 and ITGB4 protein expression levels were significantly positively correlated (Figure 8E).

TNFAIP2 expression levels positively correlated with ITGB4 in TNBC tissues

Representative IHC images of TNFAIP2 and ITGB4 protein expression in breast cancer tissues are shown. The final expression assessment was performed by combining the two scores (0–2=low, 6–7=high). A and B indicate low scores, C and D indicate high scores, and E indicates that the TNFAIP2 and ITGB4 protein expression levels are positively correlated in human TNBC specimens. Figure F is the work model of this study.

Discussion

Chemotherapies, including EPI and DDP, are the main choice for TNBC patients. Unfortunately, TNBC frequently develops resistance to chemotherapy[39]. Currently, PARP inhibitors are effective for TNBC with BRCA1/2 mutation or homologous recombination deficiency (HRD)[40–42]. PARP inhibitors can cause DNA damage repair defects and have synergistic lethal effects with HRD. Meanwhile, chemotherapy and PARP inhibitor resistance is also a major problem in the clinic.

In this study, we first found that TNFAIP2 promotes TNBC drug resistance and DNA damage repair through RAC1. Next, we found that TNFAIP2 interacts with IQGAP1and ITGB4. We verified that ITGB4 promotes TNBC drug resistance and DNA damage repair through the TNFAIP2/IQGAP1/RAC1 axis. Interestingly, we discovered for the first time that ITGB4 and TNFAIP2 promote RAC1 activity through IQGAP1. Our study reveals that ITGB4 promotes TNBC resistance through TNFAIP2-, IQGAP1-, and RAC1-mediated DNA damage repair (Figure 7). This study provides new targets for reversing TNBC resistance.

ITGB4 is well known to promote breast cancer stemness and can be activated by laminin 5[43]. In addition, ITGB4 is generally in partner with ITGA6, which is another marker of breast cancer stem cells[44] and drug resistance[43]. Therefore, whether ITGA6 has similar functions needs further study. It was reported that ITGB4 activates RAC1[45], but the mechanism is unclear. For the first time, we revealed that ITGB4 activates RAC1 through TNFAIP2 and IQGAP1. More importantly, ITGB4 promotes drug resistance through the TNFAIP2/IQGAP1/RAC1 axis.

TNFAIP2 plays important roles in different cellular and physiological processes, including cell proliferation, adhesion, migration, membrane TNT formation, angiogenesis, inflammation and tumorigenesis[14]. We previously found that TNFAIP2 was regulated by KLF5 and interacted with the small GTPases RAC1 and Cdc42, thereby regulating the actin cytoskeleton and cell morphology in breast cancer cells[9]. However, the detailed mechanism is not clearly understood. In this study, we found that IQGAP1 mediates this process. IQGAP1 is a crucial regulator of cancer development by scaffolding and facilitating different oncogenic pathways, especially RAC1/Cdc42, thus affecting proliferation, adhesion, migration, invasion, and metastasis[46]. In addition, IQGAP1 is increased during the differentiation of ovarian cancer stem cells and promotes aggressive behaviors[47]. In our study, we found that TNFAIP2 interacts with IQGAP1 and thus activates RAC1 to induce chemotherapy and PARP inhibitor drug resistance.

Furthermore, TNFAIP2 was reported to induce epithelial-to-mesenchymal transition and confer platinum resistance in urothelial cancer cells[12], and in embryonic stem cell (ESC) differentiation, TNFAIP2wasfound to be important in controlling lipid metabolism, which supports the ESC differentiation process and planarian organ maintenance[48]. Another study found that TNFAIP2 overexpression enhanced TNT-mediated autophagosome and lysosome exchange, preventing advanced glycation end product (AGE)-induced autophagy and lysosome dysfunction and apoptosis[49]. In cancer treatment, TNFAIP2 was chosen as one of the six signature genes predicting chemotherapeutic and immunotherapeutic efficacies, with high-senescore patients benefiting from immunotherapy and low-senescore patients responsive to chemotherapy[50].

These reports provide a possible explanation for previous studies showing that ITGB4 is important in EMT and cancer stemness. According to our results that there is an interaction between ITGB4 and TNFAIP2, ITGB4 might regulate EMT and stemness through TNFAIP2.TNFAIP2 is one of the important factors induced by tumor necrosis factor alpha (TNFα). Interestingly, TNFαrelease could be induced by therapeutic drugs from multiple tumor cell lines. The acquisition of docetaxel resistance was accompanied by increased constitutive production of TNFα[51]. In addition, TNFα is a key tumor-promoting effector molecule secreted by tumor-associated macrophages. In vitro neutralizing TNFα was observed to inhibit tumor progression and improve the curative effect of bevacizumab[52]. Therefore, the mechanism by which TNFα promotes chemotherapeutic resistance in breast cancer should be further investigated. For future studies, it will be important to develop Tnfaip2 knockout mice to investigate the exact role of TNFAIP2 physiologically. According to recent studies and our findings, agents targeting the interaction among ITGB4/TNFAIP2/IQGAP1 would be a promising trend for developing drugs to overcome the resistance phenomenon.

In summary, ITGB4 and TNFAIP2 play important roles in breast cancer chemoresistance. TNFAIP2 activates RAC1 to promote chemoresistance through IQGAP1. In addition, ITGB4 activates RAC1 through TNFAIP2 and IQGAP1 and confer DNA damage-related drug resistance in TNBC (Figure 8F). These results indicate that the ITGB4/TNFAIP2/IQGAP1/RAC1 axis provides potential therapeutic targets to overcome DNA damage-related drug resistance in TNBC.

Data availability statement

The authors confirm that the data supporting the findings of this study are available within the article and its supplementary materials.

Supporting information